Initial spatial conditions in simulation models: the missing leg of sensitivity analyses?

### 2017-08-04 paper.tex
.

In order to generate density grids which correspond to empirical density distributions, we select among the generated grids using an objective function which matches the point cloud of 110 metropolitan areas in Europe described by four dimensions of spatial structure : their concentration index, hierarchy index, centrality index and continuity index (cf. \cite{LeNechet2015}). A stochastic exploration of a Latin Hypercube Sampling\comment[FL]{ce n'est pas explicite} of 2000 points in the 4-dimensional space of parameters {$\alpha$, $\beta$, $n$, $N$} gives a subset of 170 interesting grids matching empirical densities, which constituted our set of different initial spatial conditions. These are further clustered into three classes of morphology: compact (e.g. Vienna), polycentric (Liege) and discontinuous (Augsburg) in order to evaluate the non-trivial effects of urban form on simulation results. We select 15 grids of each type to work with in the computation of sensitivity analysis. \comment[FL]{pourquoi 15?} \comment[FL]{dire explicitement ce qu'on va faire avec ces catégories dans la partie résultats?}

%cite{Gauvinetal2009}

%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Comparing Phase Diagrams}
\comment[FL]{j'ai proposé de mettre cette partie au début de method et de la simplifier}

%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Model Exploration Workflows as an meta-sensitivity analysis method}
\comment[FL]{item}

@@

\subsection{Experiment design}
For Schelling's model, we explore a grid of a basic parameter space of the model, which is composed of two dimensions: the minimum proportion of similar agents required in the neighbourhood for the agent to be satisfied (or intolerance level) $S\in \left[0;1\right]$ and the vacancy rate of the city $V\in \left[0;1\right]$, which then defines the number of green and red agents, present in equal proportion for this experiment. Each parameter is binned into 10 values, giving 1000 parameter points. We run 100 repetitions for each configuration, what yield reasonable convergence properties. The initial spatial configuration varies across 45 different grids, chosen among the ones which are most representative of the three types of urban morphology: 15 compact grids (like Vienna, cf figure \ref{fig:densityTypes}), 15 polycentric grids (like Liege) and 15 discontinuous grids (like Augsburg). The full experiment thus equates to 4,500,000 simulations (1000 parameter combinations x 45 density grids x 100 replications). We use OpenMOLE to distribute the computation, and apply segregation measures to characterise the results.

@@

%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/relativedistance_metaparams}\\
\includegraphics[width=0.4\textwidth]{figures/alpha_beta_schemas}\\
\caption{\textbf{Relative distances of phase diagrams by initial spatial grids described by the meta-parameters of their generator.} (Top) Relative distance as a function of meta-parameters $\alpha$ (strength of preferential attachment) and $\beta$ (strength of diffusion process). \comment[CC]{y'a un truc qui cloche avec cette figure: le nom des axes! soit alpha et beta, et on dit ce que c'est en legende, soit strength / diffusion, mais ca ne me parait pas la meilleure solution. en tout cas, pas de middle ground sinon c'est confusant, pour parler parfaitement frenglish} (Bottom) Corresponding stylised density grids }
\label{fig:sugarscape-distance-meta}
\end{figure}
%%%%%%%%%%%%%

@@

\begin{table*}[htbp]
\centering
\begin{tabular}{|l|ll|ll|ll|}
\hline
Simulation outcome by segregation index: & \multicolumn{2}{c|}{\textbf{Dissimilarity}} & \multicolumn{2}{c|}{\textbf{Entropy}} & \multicolumn{2}{c|}{\textbf{Moran's I}} \\ \hline
\textbf{Intercept}                       & -0.007 ***  & -0.03 ***                     & -0.028 ***        & -0.041 ***        & 0.127 ***        & 0.129 ***            \\ \hline
\textbf{Similarity Wanted (S)}           & -0.007 ***  & -0.007 ***                    & -0.007 ***        & -0.007 ***        & -0.005 ***       & -0.005 ***           \\ 
\textbf{Vacancy Rate (V)}                & 0.294 ***   & 0.294 ***                     & 0.204 ***         & 0.294 ***         & 0.204 ***        & \textit{0.099}       \\ \hline
\textbf{Density Grid = Polycentric}      &             & 0.038 ***                     &                   & 0.023 ***         &                  & -0.006 ***           \\ 
\textbf{Density Grid = Discontinuous}    &             & 0.03 ***                     &                   & 0.017 ***         &                  & -0.001 ***           \\ \hline
\textbf{R2 (\%)}                         & 15.01       & 15.57  & 7.62              & 7.83              & 2.04             & 2.05                 \\ 
\textbf{\# of observations (sim. runs)}  & 4,500,000   & 4,500,000                     & 4,500,000         & 4,500,000         & 4,500,000        & 4,500,000            \\ \hline
\end{tabular}
\caption{\textbf{Regression of the segregation level of Schelling simulation with order parameters and type of city grid.} *** means that the estimate is significant.}
\label{tab:regressionSchelling}
\end{table*}

>>>>

In order to generate density grids which correspond to empirical density distributions, we select among the generated grids using an objective function which matches the point cloud of 110 metropolitan areas in Europe described by four dimensions of spatial structure : their concentration index, hierarchy index, centrality index and continuity index (cf. \cite{LeNechet2015}). We sample the meta-parameter space using a Latin Hypercube Sampling, which is a convenient technique to have a point cloud with high discrepancy. We sample 2000 points in the 4-dimensional space of parameters {$\alpha$, $\beta$, $n$, $N$}. It yields a subset of 170 interesting grids matching empirical densities, which constituted our set of different initial spatial conditions. These are further clustered into three classes of morphology: compact (e.g. Vienna), polycentric (Liege) and discontinuous (Augsburg) in order to evaluate the non-trivial effects of urban form on simulation results. We select 15 grids of each type to work with in the computation of sensitivity analysis. \comment[FL]{pourquoi 15?} \comment[FL]{dire explicitement ce qu'on va faire avec ces catégories dans la partie résultats?}

@@

\subsection{Experiment design}
For Schelling's model, we explore a grid of a basic parameter space of the model, which is composed of two dimensions: the minimum proportion of similar agents required in the neighbourhood for the agent to be satisfied (or intolerance level) $S\in \left[0;1\right]$ and the vacancy rate of the city $V\in \left[0;1\right]$, which then defines the number of green and red agents, present in equal proportion for this experiment. Each parameter is binned into 10 values, giving 1000 parameter points. \comment[JR]{apparemment le sampling est sobol ou lhs pour l'espace des paramètres}
We run 100 repetitions for each configuration, what yield reasonable convergence properties. The initial spatial configuration varies across 45 different grids, chosen among the ones which are most representative of the three types of urban morphology: 15 compact grids (like Vienna, cf figure \ref{fig:densityTypes}), 15 polycentric grids (like Liege) and 15 discontinuous grids (like Augsburg). The full experiment thus equates to 4,500,000 simulations (1000 parameter combinations x 45 density grids x 100 replications). We use OpenMOLE to distribute the computation, and apply segregation measures to characterise the results.

@@

%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/relativedistance_metaparams}\\
\includegraphics[width=0.4\textwidth]{figures/alpha_beta_schemas}\\
\caption{\textbf{Relative distances of phase diagrams by initial spatial grids described by the meta-parameters of their generator.} (Top) Relative distance as a function of meta-parameters $\alpha$ (strength of preferential attachment) and $\beta$ (strength of diffusion process). (Bottom) Corresponding stylised density grids }
\label{fig:sugarscape-distance-meta}
\end{figure}
%%%%%%%%%%%%%

@@

\begin{table*}[htbp]
\caption{\textbf{Regression of the segregation level of Schelling simulation with order parameters and type of city grid.} *** means that the estimate is significant.\label{tab:regressionSchelling}}
\begin{adjustwidth}{-0.5cm}{-0.5cm}
\centering
\begin{tabular}{|m{2cm}|ll|ll|ll|}
\hline
Simulation outcome by segregation index: & \multicolumn{2}{c|}{\textbf{Dissimilarity}} & \multicolumn{2}{c|}{\textbf{Entropy}} & \multicolumn{2}{c|}{\textbf{Moran's I}} \\ \hline
\textbf{Intercept}                       & -0.007 ***  & -0.03 ***                     & -0.028 ***        & -0.041 ***        & 0.127 ***        & 0.129 ***            \\ \hline
\textbf{Similarity Wanted (S)}           & -0.007 ***  & -0.007 ***                    & -0.007 ***        & -0.007 ***        & -0.005 ***       & -0.005 ***           \\ 
\textbf{Vacancy Rate (V)}                & 0.294 ***   & 0.294 ***                     & 0.204 ***         & 0.294 ***         & 0.204 ***        & \textit{0.099}       \\ \hline
\textbf{Density Grid = Polycentric}      &             & 0.038 ***                     &                   & 0.023 ***         &                  & -0.006 ***           \\ 
\textbf{Density Grid = Discontinuous}    &             & 0.03 ***                     &                   & 0.017 ***         &                  & -0.001 ***           \\ \hline
\textbf{R2 (\%)}                         & 15.01       & 15.57  & 7.62              & 7.83              & 2.04             & 2.05                 \\ 
\textbf{\# of observations (sim. runs)}  & 4,500,000   & 4,500,000                     & 4,500,000         & 4,500,000         & 4,500,000        & 4,500,000            \\ \hline
\end{tabular}
\end{adjustwidth}
\end{table*}
